Modeling and Simulation of Load Balancing Strategies for Computing in High Energy Physics

### 2018-10-04 paper/paper.tex
Implement comments

\begin{abstract}
	The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years. To cope with this increasing load, most efficient usage of the resources is mandatory. Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system. 
	As first step to meet both challenges, a more efficient utilization of the grid and coping with the rising complexity of the system, a simulation for the WLCG should be created. \todo{Vorherigen Satz umbauen, Einschub im vorderen Teil schwer verständlich} There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.

@@

To solve both challenges, a more efficient usage of the grid and coping with the rising complexity of the system\todo{Einschub umformulieren}, a simulation for the WLCG should be created.\todo{``we aim to'' statt ``should'', auch im Abstract}
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio. This leads to the research question which modifications are needed so that Palladio can be used to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio \todo{lieber more high level: ``can we simulate?''. Eher: This paper presents prototype simulation model and extensions of Palladio?}. To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

@@

The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters. Both steps can be done via the UI of Palladio and do not require manually modeling \todo{Unklar, weil UI nach manual klingt.}. At this point the users have a valid and calibrated model of GridKa with its resource and the workload. The properties of this model are described in \Cref{sec:model}.

@@

\section{Modeling Computing Jobs and Resources}
\label{sec:model}\todo{Grafik?}
This section describes how the computing jobs and working nodes are modeled using Palladio.
The component model describes each type of CMS computing job, such as analysis or reprocessing jobs, with their resource demands. These demands are the CPU demand, the I/O demand, the number of used CPU cores, the required job slots, and the number of processed events.

@@

Each type of working node in GridKa is modeled in the resource model. Each node type has a CPU and I/O processing rate, number of cores, number of job slots, and number of instances of this node type. The types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.

@@

The usage model defines which and how many jobs are scheduled. To make sure that there is always a high load on the system, we used a closed workload with enough jobs that the system is fully utilized. Each job type has a configurable share of the overall load.

We extended Palladio with a \textit{load balancing strategy action}, so that we can simulate load balancing strategies.
This load balancing strategy action is contained in a load balancing component.
In the assembly model all job components are connected with the load balancing component, which is the entry point of the system.
This action decides on which node a job should be scheduled based on the currently free job slots of each working node. The load balancing strategies can be extended to evaluate different strategies.

\section{Model Parameter Extraction and Model Creation}
\label{sec:param}\todo{Links zu den Projekten einfügen, wo langfristig archivieren? Zendodo?}

@@

For the purposes of this proof of concept, jobs were then grouped according to their type as indicated in the job reports, but the extraction framework allows to be configured for arbitrary groupings of the extracted jobs.
Job resource requirement distributions for the different job types are extracted as distributions based on the empirical distributions in the data sets used for calibration, in the form of discretized probability density functions.

The parameter extraction tool exports model parameter sets as JSON files.
To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
This ensures an efficient and flexible model creation process.

>>>>

\begin{abstract}
	The amount of data to be processed by experiments in high energy physics (HEP) is tremendously increasing in the coming years. To cope with this increasing load, most efficient usage of the resources is mandatory. Furthermore, the computing resources for user jobs in HEP will be increasingly distributed and heterogeneous, resulting in a more difficult scheduling due the increasing complexity of the system. 
	We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
	There is currently no simulation existing helping the operators of the grid to make the correct decisions while optimizing the load balancing strategy.

@@

We aim to create a simulation for the WLCG helping the HEP community to solve both challenges: a more efficient utilization of the grid and coping with the rising complexity of the system.
Currently, there is no model or simulation in existence to evaluate the load balancing strategies of the grid, helping the operators to make the correct decisions while optimizing the load balancing strategy.
\Cref{related} describes different attempts to simulate the WLCG or parts of it. However, they showed scalability issues or the simulations do not have the necessary detail level to simulate effects on working nodes that is necessary to evaluate load balancing strategies.

@@

The contribution of this paper is the modeling and simulating of CMS computing jobs at the Tier 1 center GridKa as a proof of concept.
The model and simulation are created using the Palladio simulator, see \Cref{sec:palladio} for details on Palladio. We chose Palladio to archive a scalable simulation because it is a comparably abstract simulator (\Cref{related} gives an overview on other simulators and approaches).
This leads to the research question how we can use Palladio to evaluate load balancing strategies for the CMS computing jobs, because the simulation of load balancing strategies is a new use case for Palladio. To realise it we created an extension for the simulator that enables us to simulate load balancing strategies and furthermore, we created an automated model parameter analysis and model creation, in order to increase the usability of the simulation.
This model enables us to evaluate different load balancing strategies for GridKa and enables us to draw conclusions about the best optimizations of the load balancing for the WLCG. Furthermore, it can be used as a foundation for models and simulations of larger parts of the WLCG.

@@

The next step is the generation of the model of GridKa and the injection of the previously extracted performance parameters. Both steps do not require manually modeling. At this point the users have a valid and calibrated model of GridKa with its resource and the workload. The properties of this model are described in \Cref{sec:model}.

@@

\section{Modeling Computing Jobs and Resources}
\label{sec:model}\todo{Grafik?}
This section describes how the computing jobs and working nodes are modeled using Palladio.
The component model describes each type of CMS computing job, such as analysis or reprocessing jobs, with their resource demands. These demands are the \textit{CPU demand}, the\textit{ I/O demand}, the \textit{number of used CPU cores}, the \textit{required job slot}s, and the \textit{number of processed events}.

@@

Each type of working node in GridKa is modeled in the resource model. Each node type has a \textit{CPU and I/O processing rate}, \textit{number of cores}, \textit{number of job slots}, and \textit{number of instances} of this node type. The types of computing nodes are duplicated according to their specified number of instances as a preprocessing step of the simulation, resulting in the simulation of each working node.

@@

The usage model defines which and how many jobs are scheduled. To make sure that there is always a high load on the system, we used a closed workload with enough jobs that the system is fully utilized. Each job type has a configurable \textit{job share} of the overall load.

We extended Palladio with a \textit{load balancing strategy action}\footnote{https://github.com/PalladioSimulator/Palladio-Addon-ArchitecturalTemplates-LoadBalancing}, so that we can simulate load balancing strategies.
This load balancing strategy action is contained in a load balancing component.
In the assembly model all job components are connected with the load balancing component, which is the entry point of the system.
This action decides on which node a job should be scheduled based on the currently free job slots of each working node. The load balancing strategies can be extended to evaluate different strategies.

\section{Model Parameter Extraction and Model Creation}
\label{sec:param}

@@

For the purposes of this proof of concept, jobs were then grouped according to their type as indicated in the job reports, but the extraction framework allows to be configured for arbitrary groupings of the extracted jobs.
Job resource requirement distributions for the different job types are extracted as distributions based on the empirical distributions in the data sets used for calibration, in the form of discrete probability density functions.

The parameter extraction tool\footnote{https://github.com/mxsg/CMS-Model-Calibration} exports model parameter sets as JSON files.
To simplify calibrating simulation models with these parameter sets, we created a Palladio plugin\footnote{https://github.com/mxsg/WLCG-Model-Construction} that can be used to automatically complete a blueprint of the simulation model (which defines the structure of the model as described in section \ref{sec:model}) with arbitrary parameter sets.
This ensures an efficient and flexible model creation process.
